Bayesian regression facilitates quantitative modelling of cell metabolism

### 2023-10-31 paper/main.qmd
Update paper

To demonstrate how Maud's output can be used to yield useful metabolic
insights we used the results of our case study to explain why the flux of
the enzyme $GNMT$ is higher in dataset 1 than in dataset 12. GNMT is
an irreversible enzyme that is homotropically activated by its substrate,
competitively inhibited by its product and heterotropically inhibited by 5,10-
methylenetetrahydrofolate (mlthf). The diverse regulation makes it the ideal
test case to elucidate regulatory changes.

@fig-decomposition shows the regulatory description of GNMT, according to the
results our main case study analysis. Each curve shows the marginal posterior
distribution of the ratio of the corresponding regulatory component in dataset
1 compared with dataset 12. A positive value indicates that the component
was increased in dataset 1 relative to dataset 12, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
0 point.

>>>>

To demonstrate how Maud's output can be used to yield useful metabolic insights
we used the results of our case study to explain why the flux of the enzyme
GNMT is higher in dataset 1 than in dataset 12. GNMT is an irreversible enzyme
that is homotropically activated by its substrate, competitively inhibited by
its product and heterotropically inhibited by 5,10- methylenetetrahydrofolate
(mlthf). The diverse regulation makes it the ideal test case to elucidate
regulatory changes.

@fig-decomposition shows the regulatory description of GNMT, according to the
results our main case study analysis. Each curve shows the marginal posterior
distribution of the ratio of the corresponding regulatory component in dataset
1 compared with dataset 12. A positive value indicates that the component
was increased in dataset 1 relative to dataset 12, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
zero point.

### 2023-10-31 paper/main.tex
Update paper

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

To demonstrate how Maud's output can be used to yield useful metabolic
insights we used the results of our case study to explain why the flux
of the enzyme \(GNMT\) is higher in dataset 1 than in dataset 12. GNMT
is an irreversible enzyme that is homotropically activated by its
substrate, competitively inhibited by its product and heterotropically
inhibited by 5,10- methylenetetrahydrofolate (mlthf). The diverse
regulation makes it the ideal test case to elucidate regulatory changes.

Figure~\ref{fig-decomposition} shows the regulatory description of GNMT,
according to the results our main case study analysis. Each curve shows
the marginal posterior distribution of the ratio of the corresponding
regulatory component in dataset 1 compared with dataset 12. A positive
value indicates that the component was increased in dataset 1 relative
to dataset 12, with 0 indicating no difference. The probability,
according to our model, of the component acting in each direction is
given by the relative area under the curve on each side of the 0 point.

>>>>

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

To demonstrate how Maud's output can be used to yield useful metabolic
insights we used the results of our case study to explain why the flux
of the enzyme GNMT is higher in dataset 1 than in dataset 12. GNMT is an
irreversible enzyme that is homotropically activated by its substrate,
competitively inhibited by its product and heterotropically inhibited by
5,10- methylenetetrahydrofolate (mlthf). The diverse regulation makes it
the ideal test case to elucidate regulatory changes.

Figure~\ref{fig-decomposition} shows the regulatory description of GNMT,
according to the results our main case study analysis. Each curve shows
the marginal posterior distribution of the ratio of the corresponding
regulatory component in dataset 1 compared with dataset 12. A positive
value indicates that the component was increased in dataset 1 relative
to dataset 12, with 0 indicating no difference. The probability,
according to our model, of the component acting in each direction is
given by the relative area under the curve on each side of the zero
point.
